Allocation scores, WIS, and CRPS via decision theory

### 2023-09-28 alloscore_manuscript/methods_apg.tex
intro opportunity loss; fix WIS derivation; add copula in Norm1.R

We begin by imagining a public health worker trying to decide at time $t_0$ what quantity $x$ to procure of a protective resource such as beds, staff, or medicine in anticipation of $y$ new infections that will be observed (and create resource demand) at time $t_1>t_0$. The resource has a unit-cost of $C$, and we assign a loss of $L$ to the result that a sick individual cannot access the resource. This specifies a loss function  
\begin{align}
s_{Q,C}(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
\end{align}
where $u_+ = \max(u,0)$.  Loss functions of this general form are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error.
The underlying decision problem here consists of action and outcome sets $\mathcal{X},\mathcal{Y}$ both equal to $\mathbb{R}_+$, the non-negative reals.  (We ignore for now any discreteness of the resource.) The results set $\mathcal{R}$ is abstractly a set of over-expenditure regrets and numbers of sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{C}$ into $\mathbb{R}_+$. We also from now on make the assumption that $C < L$ since otherwise -- at least according to this loss function -- there is no ``cost-effectiveness'' in trying to prevent even a single case of unmet need. This would imply a Bayes act $x^F$ that is identically 0, telling us nothing about the value of $F$.

The Bayes act for $C \in (0,L)$ is the solution of the first order equation 
\begin{align}
0 = \dby{}{x} E_F\left[s_{Q,C}(x,y)\right] &= E_F\left[\dby{}{x}s_{Q,C}(x,y)\right] \\
&= C + LE_F\left[\dby{}{x}(y-x)_+\right] \\
&= C - LE_F\left[\mathbf{1}\{y > x\}\right] \\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
that is, the quantile $x^F = F_i^{-1}(1 - C/L) = q_{F,\tau}$ for the probability level $\tau = 1-C/L$. (That the critical point $q_{F,\tau}$ is actually a minimum follows from \eqref{eqn:q_deriv} being non-positive to its left and non-negtive to its right, and assuming as we do for now that $F$ is smooth and strictly increasing at $q_{F,\tau}$, it is a unique minimum.) The quantile solves the decision problem with respect to the forecast $F$ by being the action which best balances expected cost against expected loss among all actions available to the decision maker.

Finally, we evaluate the scoring rule $S_Q$ on $F$ when outcome $y$ realizes by measuring the loss that occurs when the action $q_{F,\tau}$ is taken:
\begin{align}
S_Q(F,y) = s_Q(q_{F,\tau}, y) = Cq_{F,\tau} + L(y- q_{F,\tau})_{+}.
\end{align}

@@

We take into account the results of actions under a set $\mathcal{C} = \{C_j\}$ of possible costs by expanding the loss function to be a weighted mean of quantile losses under certainty about the cost:
\begin{align}
s_{Q,\mathcal{C}, p_C}(x,y) = \frac{1}{m}\sum_{j=1}^{m}p_j(C_j x_j + L(x_i - y)_+). \label{eqn:sum_CL_loss}
\end{align}
Here, $p_C = \{p_j\}, j = 1,\ldots,m$ is a prior distribution on $\mathcal{C}$ and $x = (x_1,\ldots,x_m)$ is the decision vector corresponding to all possible costs.  $p_C$ would be provided by the decision maker (or perhaps a separate expert), and we emphasize that we are considering it here as part of the loss function $s_{Q,\mathcal{C}, p_C}$ that would be formulated in an effort to adapt forecast evaluation to the specifics of a public health decision problem. In practice, $p_C$ has usually been taken to be uniform, which we will show for a particular $\mathcal{C}$ (given $L$) leads to the WIS familiar in ID forecasting hubs. This might be interpreted as deferring to an uninformative prior.

We can also consider $C$ a continuous random variable with a prior density $f_C$ and accommodate this by letting $m \to \infty$ in \eqref{eqn:sum_CL_loss} 
which (with some mathematical care) converts the weighted mean into an integral
\begin{align}
s_{Q,f_C}(x,y) = \frac{1}{L}\int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}. \label{eqn:CRPS_loss}
\end{align}

@@

Accordingly, evaluation of the scoring rules $S_{Q,\mathcal{C}, p_C}$ and $S_{Q,f_C}$ on the forecast $F$ given an outcome $y$ amounts to summing the formula for $S_Q(F,y)$ across such a vector or function. That is,
\begin{align}
S_{Q,\mathcal{C}, p_C}(F,y) &= \frac{1}{m}\sum_{j=1}^{m}p_j(C_j q_{F,\tau_j} + L(q_{F,\tau_j} - y)_+) \\
S_{Q, f_C}(F,y) &= \frac{1}{L}\int_{0}^{L}(cq_{F,\tau(c)} + L(q_{F,\tau(c)} - y)_+) f_C(c)dc. \label{eqn:CRPS_cform}
\end{align}

How do the WIS and CRPS arise from this construction? Defining $\tau = 1-C/L$ as a random variable gives it the density 
$f_{\tau}(\tau) = -f_C\left(L(1-\tau)\right)/L$, 
and \eqref{eqn:CRPS_cform} becomes 
\begin{align}
S_{Q, f_C}(F,y) &= \int_{0}^{1}((1-\tau)q_{F,\tau} + (y-q_{F,\tau})_{+})f_{\tau}(\tau)d\tau \\
&= \int_{0}^{1}(\mathbf{1}\{y \leq q_{F,\tau}\}-\tau)(q_{F,\tau} - y)f_{\tau}(\tau)d\tau + \int_{0}^{1}(1-\tau)y f_{\tau}(\tau)d\tau \\
&= w\mathrm{CRPS}(F,y) + (1-E[\tau])y
\end{align}
where $w$CRPS is the weighted CRPS of \cite{gneiting2011weightedScoringRules} with quantile weight function $\nu(\tau) = \frac{1}{2}f_{\tau}(\tau)$. 
(The $(1-E[\tau])y$ that results from our loss function $s_Q$ adding a $Cy$ term to the standard quantile loss function does not effect the dependence of the scoring rule of $F$.)

Writing $s_{Q,\tau}(x,y) = s_{Q,C}(x,y)/L = (1-\tau)x + (y-x)_{+}$, we have
\begin{align}
s_{Q,\tau}(q_{F,\tau},y) + s_{Q,1-\tau}(q_{F,1-\tau},y) &= q_{F,\tau} + \tau(q_{F,1-\tau} - q_{F,\tau}) + (y - q_{F,\tau})_{+} + (y - q_{F,1-\tau})_{+} \\
&= y + \tau\left(q_{F,1-\tau} - q_{F,\tau} + \frac{1}{\tau}(q_{F,\tau} - y)_{+} + \frac{1}{\tau}(y - q_{F,1-\tau})_{+}\right) \\
&= y + \tau \mathrm{IS}_{2\tau}(F,y)
\end{align}
The scoring rule $\mathrm{IS}_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} 
and formalized by \cite{bracher2021evaluating} as a means for defining the weighted interval score 
\begin{align}
\mathrm{WIS}_{\alpha_{\{1: J\}}}(F, y)=\frac{1}{J+1 / 2}\left(w_0 |y-q_{F,.5}|+\sum_{j=1}^J w_j \operatorname{IS}_{\alpha_j}(F, y)\right).
\end{align}
So by taking $\mathcal{C} = \{(1 \pm \alpha_j)L/2 \mid j > 0\}\cup \{L/2\}$ and $p_\mathcal{C}$ symmetric across $L/2$, we get 
\begin{align}	
S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y) + y \sum_{j=0}^{J} w_j
\end{align}	
with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$. Setting $p_j=1/m$ and $\alpha_j=0.9, \ldots, 0.2,0.1,0.05,0.02, j = 1,\ldots,11$ gives the WIS currently used in 
the COVID-19 Forecast Hub.

@@

The action sets for the decision problem corresponding to a single certain cost $C$, a discrete cost distribution over $\{C_j\}, j = 1,\ldots,m$, and a continuous cost distribution $f_C$ with support on $[0,L]$ are now
\begin{align}
\mathcal{X}_{cert} &= \mathbb{R}_{+}^N \\
\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N = \{x(j):\{1,\ldots,m\} \to \mathbb{R}_{+}^N\}\\
\mathcal{X}_{cont} &= \{x(c):[0,\infty) \to \mathbb{R}_{+}^N\}.
\end{align}
For these ``vector-action'' problems we define \emph{total} versions of our loss functions
\begin{align}
s_{TQ, C}(x,y) &= \sum_{i=1}^N Cx_i + L(y_i-x_i)_+ \label{eqn:TQ_loss} \\
s_{TQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{i=1}^N \sum_{j=1}^{m}p_j(C_i x_{i,j} + L(x_{i,j} - y_i)_+)\\
s_{TQ, f_C}(x,y) &= \sum_{i=1}^N \int_{0}^{L} (cx_i(c) + L(x_i(c) - y_i)_+) f_C(c)dc.
\end{align}
Similarly as for the expansion to multiple cost contingencies---though now with respect to forecasts $F_i$ depending on the 
outcome component $y_i$---first order conditions for optimality decouple under the additive separability of the loss function so that the Bayes act $x^F$ for $F$ is just a vector of quantiles $q_{F,\tau} := (q_{F_1,\tau},\ldots,q_{F_N,\tau})$ or quantile functions when $\tau = 1 - c/L$, $c \in \mathcal{C}$ or $[0,L]$. Note that while $x^F=q_{F,\tau}$ only depends on the marginals 
$F_i$, it is indeed the Bayes act for the joint $F$ since $E_F[s_{Q,c}(x_i, y_i)] = E_{F_i}[s_{Q,c}(x_i, y_i)]$.

@@

In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
\begin{align}
s_{tQ}\left(x, y\right) = Ck + \sum_{i=1}^{N} L(y_i - x_i)_{+}.
\end{align}
The component of the Bayes act for the contingency that $K=k$ is thus independent of the realized cost $C$, motivating the definition of a new single loss function (component), the \emph{allocation loss function}
\begin{align}
s_{A}(x,y):= L\sum_{i=1}^{N} (y_i - x_i)_{+},
\end{align}
we use in deriving these Bayes acts components for all $\mathcal{X}_k$.

@@

\section{Solving the Allocation Problem}

The optimal allocation $x^{F,k}$ according to a forecast $F$ is a solution to the constrained optimization, or \emph{allocation} problem
\begin{align}
    (\mathrm{AP}) \quad \underset{x \in \mathbb{R}_{+}^N}{\mathrm{minimize}}\,\, E_{F}\left[s_A(Y,x)\right] \text{ subject to } 
    T_x = k. \label{AP}
\end{align}
The AP has a convex objective function and affine constraint, so from general convex optimization theory, a solution $x^{F,k}$ must satisfy a set of \emph{Karush-Kuhn-Tucker} (KKT) conditions. These include that for any $i$ such that $x_i^{F,K} > 0$,
\begin{align}
\frac{\partial}{\partial x_i} E_{F}\left[s_A(Y,x^{F,k})\right] &= L(F_i(x_i^{F,k})-1) = \lambda
\quad (\text{cf. \eqref{eqn:q_deriv}}) \label{eqn:x_stationary}
\end{align}
for some $\lambda$ independent of $i$. These equations follow from the assumption that when $x=x^{F,k}$, the objective $E_{F}\left[s_A(Y,x)\right]$ cannot be reduced by any small
resource shifts between locations and they provide a quantile formula for any non-zero optimal allocation level,
\begin{align}
x_i^{F,k} = F_i^{-1}(1-\lambda/L). \label{eqn:xfk_formula}
\end{align}
Another KKT condition is the constraint equation, which can now be written in terms of $\lambda$ as
\begin{align}
0 = \sum_{i \in I_+} F_{i}^{-1}(1-\lambda/L) - k, \label{eqn:lambda_stationary2}
\end{align}
where $I_+ = \{i \mid x_i^{F,k} > 0\}$.
Since convexity and compactness guarantee existence of a solution, the problem is now reduced to solving 
\eqref{eqn:lambda_stationary2} for the set 
$\{i \mid x_i^{F,k} > 0\}$ and a root $\lambda^{\star}$, in terms of which we can exrpess the Bayes act as
\begin{align}
x_i^{F,k} = 
\begin{cases}
F_i^{-1}(1-\lambda^{\star}/L) &  i \in I_+ \\
0 & \text{otherwise}.
\end{cases}
\end{align}
Note, however, that this is a more complicated task than just solving \eqref{eqn:lambda_stationary2}
for $\lambda^{\star}$ under the assumption that $I_+ = \{1,\ldots,N\}$.\footnote{Such an assumption would exclude fairly simple examples in which $F$ has support bounded away from $0$ in one coordinate but not in another.}

>>>>

We begin by imagining a public health worker trying to decide at time $t_0$ what quantity $x$ to procure of a protective resource such as beds, staff, or medicine in anticipation of $y$ new infections that will be observed (and create resource demand) at time $t_1>t_0$. The resource has a unit-cost of $C$, and we assign a loss of $L$ to the result that a sick individual cannot access the resource. We assume that $C < L$ since otherwise -- at least according to the quantification $L$ -- there is no ``cost-effectiveness'' in trying to prevent even a single case of unmet need.  With $C$ and $L$, there are two natural loss function we can define.  First, we have the \emph{absolute} loss  
\begin{align}
s_{Q,C,a}(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
\end{align}
where $u_+ = \max(u,0)$.  Alternatively, we have the \emph{opportunity} loss 
\begin{align}
s_{Q,C,o}(x,y) &= Cx + L(y-x)_+ - Cy \\
&= C(x-y)_+ + (L-C)(y-x)_+ \\
&= (C - L \mathbf{1}\{y > x\})(x-y) \\
&= L(\mathbf{1}\{y \leq x\} - 1 + C/L)(x - y)\label{eqn:quantile_loss_opp}
\end{align}
which considers only the loss relative to that of an oracle that is able to minimize loss \textit{ex-post} by procuring 
exactly $y$ units of the resource.

Loss functions of these general forms are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error. The opportunity loss $s_{Q,C,o}$ is sometimes called ``pinball'' loss. It has the convenien{}t 
{}property that $s_{Q,C,o}(z,z)=0$, but loses some of the interpretability of $s_{Q,C}$ (see \ref{} below).
The underlying decision problem here consists of action and outcome sets $\mathcal{X},\mathcal{Y}$ both equal to $\mathbb{R}_+$, the non-negative reals.  (We ignore for now any discreteness of the resource.) The results set $\mathcal{R}$ is abstractly a set of transfers of a monetary or similar nature and sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{R}$ into $\mathbb{R}_+$.

Note that the absolute and opportunity losses have the same dependence on action variable $x$, and therefore the same Bayes act. For $C \in (0,L)$ this is the solution of the first order equation 
\begin{align}{}
0 = \dby{}{x} E_F\left[s_{Q,C}(x,y)\right] &= E_F\left[\dby{}{x}s_{Q,C}(x,y)\right] \\
&= C + LE_F\left[\dby{}{x}(y-x)_+\right] \\
&= C - LE_F\left[\mathbf{1}\{y > x\}\right] \\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
that is, the quantile $x^F = F_i^{-1}(1 - C/L) = q_{F,\tau}$ for the probability level $\tau = 1-C/L$. (That the critical point $q_{F,\tau}$ is actually a minimum follows from \eqref{eqn:q_deriv} being non-positive to its left and non-negtive to its right, and assuming as we do for now that $F$ is smooth and strictly increasing at $q_{F,\tau}$, it is a unique minimum.) The quantile solves the decision problem with respect to the forecast $F$ by being the action which best balances expected cost against expected loss among all actions available to the decision maker.

Finally, we evaluate the scoring rules $S_{Q,C,a}$ and $S_{Q,C,o}$ on $F$ when outcome $y$ realizes by measuring the loss that occurs when the action $q_{F,\tau}$ is taken:
\begin{align}
S_{Q,C,a}(F,y) &= s_{Q,C,a}(q_{F,\tau}, y) = Cq_{F,\tau} + L(y- q_{F,\tau})_{+} \\
S_{Q,C,o}(F,y) &= s_{Q,C,o}(q_{F,\tau}, y) = (C - L \mathbf{1}\{y > x\})(q_{F,\tau}-y).
\end{align}

@@

We take into account the results of actions under a set $\mathcal{C} = \{C_j\}$ of possible costs by expanding 
(either the absolute or opportunity version of) the loss function to be a weighted mean of quantile losses under certainty about the cost:
\begin{align}
s_{Q,\mathcal{C}, p_C}(x,y) = \sum_{j=1}^{m}p_js_{Q,C_j}(x_j,y). \label{eqn:sum_CL_loss}
\end{align}
Here, $p_C = \{p_j\}, j = 1,\ldots,m$ is a prior distribution on $\mathcal{C}$ and $x = (x_1,\ldots,x_m)$ is the decision vector corresponding to all possible costs.  $p_C$ would be provided by the decision maker (or perhaps a separate expert), and we emphasize that we are considering it here as part of the loss function $s_{Q,\mathcal{C}, p_C}$ that would be formulated in an effort to adapt forecast evaluation to the specifics of a public health decision problem. In practice, $p_C$ has usually been taken to be uniform, which we will show for a particular $\mathcal{C}$ (given $L$) leads to the WIS familiar in ID forecasting hubs. This might be interpreted as deferring to an uninformative prior.

We can also consider $C$ a continuous random variable with a prior density $f_C$ and accommodate this by letting $m \to \infty$ in \eqref{eqn:sum_CL_loss} 
which (with some mathematical care) converts the weighted mean into an integral
\begin{align}
s_{Q,f_C}(x,y) = \int_{0}^{L} s_{Q,c}(x(c),y) f_C(c)dc{}. \label{eqn:CRPS_loss}
\end{align}

@@

Accordingly, evaluation of the scoring rules $S_{Q,\mathcal{C}, p_C}$ and $S_{Q,f_C}$ on the forecast $F$ given an outcome $y$ amounts to summing the formula for $S_Q(F,y)$ across such a vector or function. That is,
\begin{align}
S_{Q,\mathcal{C}, p_C}(F,y) &= \sum_{j=1}^{m}p_js_{Q,C_j}(q_{F,\tau(C_j)},y) \\
S_{Q, f_C}(F,y) &= \int_{0}^{L} s_{Q,c}(q_{F,\tau(c)},y) f_C(c)dc. \label{eqn:CRPS_cform}
\end{align}

How do the WIS and CRPS arise from this construction? Defining $\tau = 1-C/L$ as a random variable gives it the density 
$f_{\tau}(\tau) = f_C\left(L(1-\tau)\right)\cdot L$, 
and \eqref{eqn:CRPS_cform} with the opportunity loss \eqref{eqn:quantile_loss_opp} becomes 
\begin{align} 
S_{Q, f_C}(F,y) &= \int_{0}^{L}L(\mathbf{1}\{y \leq q_{F,\tau(c)}\}-\tau(c))(q_{F,\tau(c)} - y)f_{C}(c)dc \\
&= \int_{0}^{1}(\mathbf{1}\{y \leq q_{F,\tau}\}-\tau)(q_{F,\tau} - y)f_{\tau}(\tau)d\tau \\
&= w\mathrm{CRPS}(F,y)
\end{align}
where $w$CRPS is the weighted CRPS of \cite{gneiting2011weightedScoringRules} with quantile weight function $\nu(\tau) = \frac{1}{2}f_{\tau}(\tau)$. 
Using the absolute loss instead of opportunity loss would add a term $E[C]y$. This is the value of the absolute loss scoring rule 
at the oracle's point mass forecast: $S_{Q, f_C, a}(\delta_y,y)$.

To connect with the WIS, we recall it's definition by \cite{bracher2021evaluating}:
\begin{align}
\mathrm{WIS}_{\alpha_{\{1: J\}}}(F, y)=\frac{1}{J+1 / 2}\left(w_0 |y-q_{F,.5}|+\sum_{j=1}^J w_j \operatorname{IS}_{\alpha_j}(F, y)\right).
\end{align}
The scoring rule $\mathrm{IS}_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} 
and formalized by \cite{bracher2021evaluating} as
\begin{align}
\operatorname{IS}_{\alpha_j}(F, y) = \frac{\alpha_j}{2}(q_{F,\alpha_j/2}-y)_+ + \frac{\alpha_j}{2}(y-q_{F,1-\alpha_j/2})_+ 
 + (q_{F,1-\alpha_j/2} - q_{F,\alpha_j/2})
\end{align}
Thus the $\alpha_j$ are twice the probability level of the lower endpoint of a central
$1-\alpha_j$ prediction interval. The COVID-19 Forecast Hub chooses $\alpha_j=0.02, 0.05, 0.1, \ldots, 0.9, j = 1,\ldots,11$.

Now write $s_{Q,\tau, o}(x,y) = s_{Q,C, o}(x,y)/L = (1-\tau)(x-y)_+ + \tau(y-x)_+$, 
and assume $\tau < 1/2$.  We have
\begin{align}
s_{Q,\tau,o}(q_{F,\tau},y) &+ s_{Q,1-\tau,o}(q_{F,1-\tau},y) \\
&= (1-\tau)(q_{F,\tau}-y)_+ + \tau(y-q_{F,\tau})_+ 
+ \tau(q_{F,1-\tau}-y)_+ + (1-\tau)(y-q_{F,1-\tau})_+ \\ 
&= (q_{F,\tau}-y)_+ + (y-q_{F,1-\tau})_+ \\
&\quad + \tau \left[(y-q_{F,\tau})_+ - (q_{F,\tau}-y)_+ + (q_{F,1-\tau}-y)_+ - (y-q_{F,1-\tau})_+\right] \\
&= (q_{F,\tau}-y)_+ + (y-q_{F,1-\tau})_+  + \tau(q_{F,1-\tau} - q_{F,\tau}) \\
&= \tau \mathrm{IS}_{2\tau}(F,y). \label{eqn:pinball_IS}
\end{align}

To reproduce the WIS we set $\mathcal{C} = \{C_k\}, k = 0,\ldots 2J$ where 
\begin{align}
C_0 &= L/2 \\
\alpha_k/2 &= 1-C_k/L, k=1,..J \\
1-\alpha_{k-J}/2 &= C_k/L, k = J+1,\ldots,2J
\end{align}
(note that $C_k$ decrease as $k$ increases for $k>0$)
and let $p_{\mathcal{C}}$ be a probability distribution symmetric across $L/2$. This gives us
\begin{align}	
S_{Q,\mathcal{C}, p_C, o}(F,y) &= p_0 s_{Q,L/2, o}(q_{F,.5},y) + \sum_{k = 1}^{2J} p_k s_{Q,C_k, o}(q_{F, 1-C_k/L},y) \\
& =  p_0 s_{Q,L/2, o}(q_{F,.5},y) + \sum_{k = 1}^{J} L(1-C_k/L)p_k \mathrm{IS}_{2(1-C_k/L)}(F,y) \\
& = \frac{L}{2} \left(p_0|y-q_{F,.5}| + \sum_{j = 1}^{J} \alpha_j p_j \mathrm{IS}_{\alpha_j}(F,y)\right) \\
& \propto \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y) \label{eqn:WIS_propto}
\end{align}	
if we set $p_j \propto w_j/\alpha_j$ and define $\alpha_0 = 1$. The COVID-19 Forecast Hub chooses $w_j = \alpha_j/2$ 
so that the WIS there corresponds to a uniform distribution over the $C_k$ where each cost has probability
\begin{align}
p = \frac{1/2}{\sum_{k=0}^{2J} 1/2} = \frac{1/2}{J+1/2}.
\end{align}
In this case the constant of proportionality in \eqref{eqn:WIS_propto} is $L/2$.

@@

The action sets for the decision problem corresponding to a single certain cost $C$, a discrete cost distribution over $\{C_j\}, j = 1,\ldots,m$, and a continuous cost distribution $f_C$ with support on $[0,L]$ are now
\begin{align}
\mathcal{X}_{cert} &= \mathbb{R}_{+}^N \\
\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N = \{x(j):\{1,\ldots,m\} \to \mathbb{R}_{+}^N\}\\
\mathcal{X}_{cont} &= \{x(c):[0,\infty) \to \mathbb{R}_{+}^N\}.
\end{align}
For these ``vector-action'' problems we define \emph{total} versions of our loss functions
\begin{align}
s_{TQ, C}(x,y) &= \sum_{i=1}^N Cx_i + L(y_i-x_i)_+ \label{eqn:TQ_loss} \\
s_{TQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{i=1}^N \sum_{j=1}^{m}p_j(C_i x_{i,j} + L(x_{i,j} - y_i)_+)\\
s_{TQ, f_C}(x,y) &= \sum_{i=1}^N \int_{0}^{L} (cx_i(c) + L(x_i(c) - y_i)_+) f_C(c)dc.
\end{align}
Similarly as for the expansion to multiple cost contingencies---though now with respect to forecasts $F_i$ depending on the 
index $i$ of the relevant outcome component $y_i$---first order conditions for optimality decouple under the additive separability of the loss function so that the Bayes act $x^F$ for $F$ is just a vector of quantiles $q_{F,\tau} := (q_{F_1,\tau},\ldots,q_{F_N,\tau})$ or quantile functions when $\tau = 1 - c/L$, $c \in \mathcal{C}$ or $[0,L]$. Note that while $x^F=q_{F,\tau}$ only depends on the marginals 
$F_i$, it is indeed the Bayes act for the joint $F$ since $E_F[s_{TQ,c}(x, y)] = \sum_{i=1}^{N}E_{F_i}[s_{Q,c}(x_i, y_i)]$.

@@

In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
\begin{align}
s_{TQ}\left(x, y\right) = Ck + \sum_{i=1}^{N} L(y_i - x_i)_{+}.
\end{align}
The component $x_i^{F,k}$ of the Bayes act for the contingency that $K=k$ is thus independent of the realized cost $C$, motivating the definition of a new single loss function (component), the \emph{allocation loss function}
\begin{align}
s_{A}(x,y):= L\sum_{i=1}^{N} (y_i - x_i)_{+},
\end{align}
we use in deriving these Bayes acts components for all $\mathcal{X}_k$.

We show in section \ref{sec:solvingAP} that $x^{F,k}$ is defined by a single probability level $\tau$ just as was the Bayes act $x^{F,c}$ for $s_{TQ,C}$ on $\mathcal{X}_{cert}$, but that in this case, $\tau = \tau(k, F)$ depends not only on the decision problem 
parameter $k$, but also on the forecast $F$ itself.

@@

\section{Solving the Allocation Problem}
\label{sec:solvingAP}

The optimal allocation $x^{F,k}$ according to a forecast $F$ is a solution to the constrained optimization, or \emph{allocation} problem
\begin{align}
    (\mathrm{AP}) \quad \underset{x \in \mathbb{R}_{+}^N}{\mathrm{minimize}}\,\, E_{F}\left[s_A(Y,x)\right] \text{ subject to } 
    T_x = k. \label{AP}
\end{align}
The AP has a convex objective function and affine constraint, so from general convex optimization theory, a solution $x^{F,k}$ must satisfy a set of \emph{Karush-Kuhn-Tucker} (KKT) conditions. These include that for any $i$ such that $x_i^{F,K} > 0$,
\begin{align}
\frac{\partial}{\partial x_i} E_{F}\left[s_A(Y,x^{F,k})\right] &= L(F_i(x_i^{F,k})-1) = \lambda
\quad (\text{cf. \eqref{eqn:q_deriv}}) \label{eqn:x_stationary}
\end{align}
for some $\lambda$ independent of $i$. These equations follow from the assumption that when $x=x^{F,k}$, the objective $E_{F}\left[s_A(Y,x)\right]$ cannot be reduced by any small
resource shifts between locations and they provide a quantile formula for any non-zero optimal allocation level,
\begin{align}
x_i^{F,k} = F_i^{-1}(1-\lambda/L). \label{eqn:xfk_formula}
\end{align}
Another KKT condition is the constraint equation, which can now be written in terms of $\lambda$ as
\begin{align}
0 = \sum_{i \in I_+} F_{i}^{-1}(1-\lambda/L) - k, \label{eqn:lambda_stationary2}
\end{align}
where $I_+ = \{i \mid x_i^{F,k} > 0\}$.
Since convexity and compactness guarantee existence of a solution, the problem is now reduced to solving 
\eqref{eqn:lambda_stationary2} for the set 
$\{i \mid x_i^{F,k} > 0\}$ and a root $\lambda^{\star}$, in terms of which we can express the Bayes act as
\begin{align}
x_i^{F,k} = 
\begin{cases}
F_i^{-1}(1-\lambda^{\star}/L) &  i \in I_+ \\
0 & \text{otherwise}.
\end{cases}
\end{align}
Note, however, that this is a more complicated task than just solving \eqref{eqn:lambda_stationary2}
for $\lambda^{\star}$ under the assumption that $I_+ = \{1,\ldots,N\}$.\footnote{Such an assumption would exclude fairly simple examples in which the marginal densities $f_i$ of $F$ have different non-zero limits as the $x_i \to 0$.}

The task can be simplified though by turning our attention to the \emph{dual} formulation of the AP. (See e.g. \cite{ruszczynski2011nonlinear}, Chapter 4, which we follow here.) This involves recasting the KKT equations \eqref{eqn:x_stationary}
and the constraint $T_x=k$ as the first order conditions in variables $x$ and $\lambda$, respectively, for $(x^{F,k}, \lambda^{\star})$ to be a saddle point of the \emph{Lagrangian}
\begin{align}{}
\mathcal{L}(x, \lambda) = \mathcal{L}(x, \lambda; F, k, L) 
&= E_F\left[s_A(Y,x)\right] + \lambda(T_x - k) \\
&= L\sum_{i=1}^{N} E_{F_i}[(Y_i - x_i)_{+}] + \lambda\left(\sum_{i=1}^{N} x_i - k\right).  
\end{align}
in the region $\{x \geq 0, \lambda \in \mathbb{R}\}$.\footnote{The Lagrangian in this context has the interpretation as the best (lowest) objective function value we can achieve if we forced to choose $x$ but can buy or get credit for changes in the constraint level $k$ 
at the ``price'' $\lambda$.}
The \emph{dual function} is then defined as $\mathcal{L}_D(\lambda) := \inf_{x \in \mathbb{R}_{+}^N} \mathcal{L}(x, \lambda)$. 
From general theory, $\mathcal{L}_D(\lambda)$ is concave and $\lambda^{\star}$ is its unique maximizer.
The reason why the dual approach is especially suitable in our situation is that 
our original objective function $E_F\left[s_A(Y,x)\right]$ is decomposable in the coodinates $x_i$, allowing the infimum
$\mathcal{L}_D$ to be computed as a sum of separate infima in these coordinates:
\begin{align}
\mathcal{L}_D(\lambda) &= \inf_{x \in \mathbb{R}_{+}^N} L\sum_{i=1}^{N} E_{F_i}[(Y_i - x_i)_{+}] + \lambda\left(\sum_{i=1}^{N} x_i - k\right) \\
&= - \lambda k + \sum_{i=1}^{N} \inf_{x_i \in \mathbb{R}_{+}}\left\{ L E_{F_i}[(Y_i - x_i)_{+}] + \lambda x_i\right\}  \label{eqn:dual_decomp} \\
&= - \lambda k + \sum_{i=1}^{N}  LE_{F_i}[(Y_i - x_i(\lambda))_{+}] + \lambda x_i(\lambda).
\end{align}
Here, $x_i(\lambda)$ is any member of the set 
\begin{align}
\chi_i(\lambda) = \underset{x_i \in \mathbb{R}_{+}}{\mathrm{arginf}} \left\{ L E_{F_i}[(Y_i - x_i)_{+}] + \lambda x_i\right\}
\end{align}
which for $F_i$ in a standard parametric family will often simply be $\{F_i^{-1}(1-\lambda/L)\}$ but could an interval containing $0$ or $\infty$.
While $\mathcal{L}_D(\lambda)$ may not be differentiable when the $x_i$'s attaining the infima in \eqref{eqn:dual_decomp} are zero, it does have as a well-defined subdifferential which is the sum of intervals
\begin{align}
\partial \mathcal{L}_D(\lambda) 
&= \sum_{i=1}^N \chi_i(\lambda) - k \\
&= \left[ \sum_{i=1}^N \min(\chi_i(\lambda))  - k, \sum_{i=1}^N \max(\chi_i(\lambda))  - k\right]
\end{align}
